### 2023-08-21 sub/smimr/2/cover.tex

\documentclass{uiletter}
\usepackage{hyperref}
\hypersetup{colorlinks=true, citecolor=black, urlcolor=cyan}

\begin{document}

\StartLetter{}{Dear editors of {\em Statistical Methods in Medical Research},}

We are submitting a revised version of the manuscript ``Cross validation approaches for penalized Cox regression'' for your consideration to be published in \emph{Statistical Methods in Medical Research}.

This version has been revised and improved in several ways, most importantly by carrying out a more thorough literature search with respect to the deviance residual method. In our original submission, we pointed out that this was an approach one could try, but that it didn't work very well. This was a bit confusing, as it seemed like we were proposing an approach that didn't work very well (in addition to one that did). Since the original submission, we were able to find a number of other papers in which this approach has indeed been utilized, which makes our investigation of this method more clearly relevant.

In addition, we incorporated all the excellent feedback we received from the reviewers, and feel that we have addressed all their concerns in this revision.

We realize that we are quite late with this resubmission -- my coauthor Biyue moved to a new institution and due to a miscommunication, we each thought the other had resubmitted the paper. I communicated with Prof. Chris Metcalfe, who encouraged us to resubmit the paper despite this on August 12, 2023. Thank you very much for your consideration of this revised manuscript.

\EndLetter

\end{document}