### 2024-01-11 sub/smimr/3/cover.tex
Formatting for SMIMR resubmission

\documentclass{uiletter}
\usepackage{hyperref}
\hypersetup{colorlinks=true, citecolor=black, urlcolor=cyan}

\begin{document}

\StartLetter{}{Dear editors of {\em Statistical Methods in Medical Research},}

We are submitting another revision of the manuscript ``Cross-validation approaches for penalized Cox regression'' for your consideration to be published in \emph{Statistical Methods in Medical Research}.

The revisions here are relatively modest -- both reviewers were happy with the paper, although reviewer \#1 pointed out some places where clarifications that could be made and inconsistencies addressed. These have all been addressed now.

Thank you very much for your consideration of this revised manuscript.

\EndLetter

\end{document}